---
title: "RF_Classification_BDvsHD"
author: "Leah Guthrie"
date: "12/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("caret")
library("randomForest")
```


# Importing Data
```{r}
ASV <- read.csv("../data/MISO-fecal-asv-long.csv")
metadata <- read.csv("../data/MISO-metadata.csv")
urineMetabolome <- read.csv("../data/MISO-urine-metabolites-long.csv")
plasmaMetabolome <- read.csv("../data/MISO-plasma-metabolites-long.csv")
fecalMetabolome <- read.csv("../data/MISO-fecal-metabolites-long.csv")
```

# Plasma Metabolome

## Filter by time point - center and scale data
```{r}
plasmaMetabolomet2vst3 <- plasmaMetabolome %>% filter(timepoint != 5) %>% filter(timepoint != 1) %>% filter(timepoint != 4)
plasmaMetabolomet2vst3x <- plasmaMetabolomet2vst3 %>% select(-c(X,ID2,subject,timepoint,timepointgroup,phase)) %>%
  mutate_if(is.numeric, scale)
plasmaMetabolomet2vst3x
plasmaMetabolomet2vst3y <- plasmaMetabolomet2vst3 %>% select(timepointgroup)
plasmaMetabolomet2vst3y$timepointgroup <- gsub('BD', 0, plasmaMetabolomet2vst3y$timepointgroup)
plasmaMetabolomet2vst3y$timepointgroup <- gsub('HD', 1, plasmaMetabolomet2vst3y$timepointgroup)
plasmaMetabolomet2vst3y
```

## Training: 80%; Test: 20%
```{r}
x <- plasmaMetabolomet2vst3x
y <- factor(plasmaMetabolomet2vst3y$timepointgroup)

set.seed(2021)
inTrain <- createDataPartition(y, p = .80, list = FALSE)[,1]

x_train <- x[ inTrain, ]
x_test  <- x[-inTrain, ]

y_train <- y[ inTrain]
y_test  <- y[-inTrain]
```

##
```{r}
fsa_rfecontrol <- rfeControl(functions = rfFuncs,
                      method = "repeatedcv", 
                      repeats = 5, 
                      number = 10) 
```

## Recursive feture elimination algorithm
```{r}
# Run RFE
result_rfe1 <- rfe(x = x_train, 
                   y = y_train, 
                   sizes = c(1:53),
                   rfeControl = fsa_rfecontrol)
```


## Confusion Matrix
```{r}
plasmaConfusionMatrix <- as.data.frame(result_rfe1$fit$confusion)
write.csv(plasmaConfusionMatrix, "ConfusionMatrix/plasmaConfusionMatrix.csv")
write.csv(result_rfe1$results, "ConfusionMatrix/plasmaAccuracyCohen_estimates.csv")
```


# Urine Metabolome
```{r}
urineMetabolomet2vst3 <- urineMetabolome %>% filter(timepoint != 5) %>% filter(timepoint != 1) %>% filter(timepoint != 4)
urineMetabolomet2vst3x <- urineMetabolomet2vst3 %>% select(-c(ID2,subject,timepoint,timepointgroup,phase))
urineMetabolomet2vst3x <- urineMetabolomet2vst3x %>%
  mutate_if(is.numeric, scale)
urineMetabolomet2vst3y <- urineMetabolomet2vst3 %>% select(timepointgroup)
urineMetabolomet2vst3y$timepointgroup <- gsub('BD', 0, urineMetabolomet2vst3y$timepointgroup)
urineMetabolomet2vst3y$timepointgroup <- gsub('HD', 1, urineMetabolomet2vst3y$timepointgroup)
```


## Training: 80%; Test: 20%
```{r}
x.u <- urineMetabolomet2vst3x
y.u <- factor(urineMetabolomet2vst3y$timepointgroup)

set.seed(2021)
inTrain.u <- createDataPartition(y.u, p = .80, list = FALSE)[,1]

x_train.u <- x.u[ inTrain.u, ]
x_test.u  <- x.u[-inTrain.u, ]

y_train.u <- y.u[ inTrain.u]
y_test.u  <- y.u[-inTrain.u]
```


## Recursive feture elimination algorithm
```{r}
result_rfe.u <- rfe(x = x_train.u, 
                   y = y_train.u, 
                   sizes = c(1:25),
                   rfeControl = fsa_rfecontrol)
```


```{r}
urineConfusionMatrix <- as.data.frame(result_rfe.u$fit$confusion)
write.csv(urineConfusionMatrix, "ConfusionMatrix/urineConfusionMatrix.csv")
write.csv(result_rfe.u$results, "ConfusionMatrix/urineAccuracyCohen_estimates.csv")
```


# Fecal metabolome

##
```{r}
fecalMetabolomet2vst3 <- fecalMetabolome %>% filter(timepoint != 5) %>% filter(timepoint != 1) %>% filter(timepoint != 4)
fecalMetabolomet2vst3x <- fecalMetabolomet2vst3 %>% select(-c(X,ID2,subject,timepoint,timepointgroup,phase)) %>%
  mutate_if(is.numeric, scale)
fecalMetabolomet2vst3x
fecalMetabolomet2vst3y <- fecalMetabolomet2vst3 %>% select(timepointgroup)
fecalMetabolomet2vst3y$timepointgroup <- gsub('BD', 0, fecalMetabolomet2vst3y$timepointgroup)
fecalMetabolomet2vst3y$timepointgroup <- gsub('HD', 1, fecalMetabolomet2vst3y$timepointgroup)
fecalMetabolomet2vst3y
```


# Training: 80%; Test: 20%
```{r}
x.f <- fecalMetabolomet2vst3x
y.f <- factor(fecalMetabolomet2vst3y$timepointgroup)

set.seed(2021)
inTrain.f <- createDataPartition(y.f, p = .80, list = FALSE)[,1]

x_train.f <- x.f[ inTrain.f, ]
x_test.f  <- x.f[-inTrain.f, ]

y_train.f <- y.f[ inTrain.f]
y_test.f  <- y.f[-inTrain.f]
```

## Recursive feture elimination algorithm
```{r}
result_rfe.f <- rfe(x = x_train.f, 
                   y = y_train.f, 
                   sizes = c(1:320),
                   rfeControl = fsa_rfecontrol)
```


```{r}
fecalConfusionMatrix <- as.data.frame(result_rfe.f$fit$confusion)
write.csv(fecalConfusionMatrix, "ConfusionMatrix/fecalConfusionMatrix.csv")
write.csv(result_rfe.f$results, "ConfusionMatrix/fecalAccuracyCohen_estimates.csv")
```

# Macronutrients
```{r}
macros <- metadata %>% select(c(ID2,timepoint,timepointgroup,Prot_g,Carb_g,Fat_g,TotFib_g))
macrosMetabolomet2vst3 <- macros %>% filter(timepoint != 5) %>% filter(timepoint != 1) %>% filter(timepoint != 4)
macrosMetabolomet2vst3x <- macrosMetabolomet2vst3 %>% select(-c(ID2,timepoint,timepointgroup)) %>%
  mutate_if(is.numeric, scale)
macrosMetabolomet2vst3x
macrosMetabolomet2vst3y <- macrosMetabolomet2vst3 %>% select(timepointgroup)
macrosMetabolomet2vst3y$timepointgroup <- gsub('BD', 0, macrosMetabolomet2vst3y$timepointgroup)
macrosMetabolomet2vst3y$timepointgroup <- gsub('HD', 1, macrosMetabolomet2vst3y$timepointgroup)
macrosMetabolomet2vst3y
```

# Training: 80%; Test: 20%
```{r}
x.macr <- macrosMetabolomet2vst3x
y.macr <- factor(macrosMetabolomet2vst3y$timepointgroup)

set.seed(2021)
inTrain.macr <- createDataPartition(y.macr, p = .80, list = FALSE)[,1]

x_train.macr <- x.macr[ inTrain.macr, ]
x_test.macr  <- x.macr[-inTrain.macr, ]

y_train.macr <- y.macr[ inTrain.macr]
y_test.macr  <- y.macr[-inTrain.macr]
```

## Recursive feture elimination algorithm
```{r}
result_rfe.macr <- rfe(x = x_train.macr, 
                   y = y_train.macr, 
                   sizes = c(1:4),
                   rfeControl = fsa_rfecontrol)
```


```{r}
macrosConfusionMatrix <- as.data.frame(result_rfe.macr$fit$confusion)
write.csv(macrosConfusionMatrix, "ConfusionMatrix/macrosConfusionMatrix.csv")
write.csv(result_rfe.macr$results, "ConfusionMatrix/macrosAccuracyCohen_estimates.csv")
```

# ASVs


##
```{r}
ASVt2vst3 <- ASV %>% filter(timepoint != 5) %>% filter(timepoint != 1) %>% filter(timepoint != 4)
ASVt2vst3x <- ASVt2vst3 %>% select(-c(X,ID2,subject,timepoint,timepointgroup,phase)) %>%
  mutate_if(is.numeric, scale)

ASVt2vst3x <- ASVt2vst3x %>%
    select_if(~ !any(is.na(.)))
ASVt2vst3y <- ASVt2vst3 %>% select(timepointgroup)
ASVt2vst3y$timepointgroup <- gsub('BD', 0, ASVt2vst3y$timepointgroup)
ASVt2vst3y$timepointgroup <- gsub('HD', 1, ASVt2vst3y$timepointgroup)
ASVt2vst3y
ASVt2vst3x
```


```{r}
x.asv <- ASVt2vst3x
y.asv <- factor(ASVt2vst3y$timepointgroup)

set.seed(2021)
inTrain.asv <- createDataPartition(y.asv, p = .80, list = FALSE)[,1]

x_train.asv <- x.asv[ inTrain.asv, ]
x_test.asv  <- x.asv[-inTrain.asv, ]

y_train.asv <- y.asv[ inTrain.asv]
y_test.asv  <- y.asv[-inTrain.asv]
```

## Recursive feture elimination algorithm
```{r}
result_rfe.asv <- rfe(x = x_train.asv, 
                   y = y_train.asv, 
                   sizes = c(1:500),
                   rfeControl = fsa_rfecontrol)
```


```{r}
asvConfusionMatrix <- as.data.frame(result_rfe.asv$fit$confusion)
write.csv(asvConfusionMatrix, "ConfusionMatrix/asvconfusionMatrix.csv")
write.csv(result_rfe.asv$results, "ConfusionMatrix/asvAccuracyCohen_estimates.csv")
```



## Metadata
```{r}
metadatat2vst3 <- metadata %>% filter(timepoint != 5) %>% filter(timepoint != 1) %>% filter(timepoint != 4)
metadatat2vst3 <- metadatat2vst3 %>% select(ID2,timepoint,timepointgroup,Age,BSA,Gender)
metadatat2vst3
```


```{r}
metadatat2vst3x <- metadatat2vst3 %>% select(-c(ID2,timepoint,timepointgroup))
metadatat2vst3x$Gender <- gsub('M', 0, metadatat2vst3x$Gender)
metadatat2vst3x$Gender <- gsub('F', 1, metadatat2vst3x$Gender)
metadatat2vst3x <- metadatat2vst3x %>% mutate_at(c("Gender"),as.factor) %>%
  # Center and scale numeric features
  mutate_if(is.numeric, scale)
metadatat2vst3x
```

```{r}
metadatat2vst3y <- metadatat2vst3 %>% select(timepointgroup)
metadatat2vst3y$timepointgroup <- gsub('BD', 0, metadatat2vst3y$timepointgroup)
metadatat2vst3y$timepointgroup <- gsub('HD', 1, metadatat2vst3y$timepointgroup)
metadatat2vst3y
```


```{r}
x.meta <-metadatat2vst3x
y.meta <- factor(metadatat2vst3y$timepointgroup)

set.seed(2021)
inTrain.meta <- createDataPartition(y.meta, p = .80, list = FALSE)[,1]

x_train.meta <- x.meta[ inTrain.meta, ]
x_test.meta  <- x.meta[-inTrain.meta, ]

y_train.meta <- y.meta[ inTrain.meta]
y_test.meta  <- y.meta[-inTrain.meta]
```

## Recursive feture elimination algorithm
```{r}
result_rfe.meta <- rfe(x = x_train.meta, 
                   y = y_train.meta, 
                   sizes = c(1:500),
                   rfeControl = fsa_rfecontrol)
```

```{r}
metapAccuracy <- ggplot(data = result_rfe.meta, metric = "Accuracy") + theme_bw() + panel_border(color="black")
metapKappa <- ggplot(data = result_rfe.meta, metric = "Kappa") + theme_bw() + panel_border(color="black")
metapAccuracy
metapKappa
```

```{r}
metaConfusionMatrix <- as.data.frame(result_rfe.meta$fit$confusion)
write.csv(metaConfusionMatrix, "ConfusionMatrix/metaconfusionMatrix.csv")
write.csv(result_rfe.meta$results, "ConfusionMatrix/metaAccuracyCohen_estimates.csv")
```

```{r}
save(metapAccuracy,metapKappa,
     file = "Plots/RFE-plots-clinc.Rdata")
```

```{r}
save(plasmapAccuracy,plasmapKappa, urinepAccuracy, urinepKappa, file = "Plots/RFE-plots.Rdata")
```

```{r}
save(macrospAccuracy,macrospKappa,
     file = "Plots/RFE-plots-macros.Rdata")
```

```{r}
save(asvpAccuracy,asvpKappa,
     file = "Plots/RFE-plots-asv.Rdata")
```